---
title: "HW4"
output: html_document
---
```{r, include=FALSE}
library(tidyverse)
library(rsample)
```

# 1. Section 5.4 Question 5 Data Cleaning
```{r}
dat <- read.csv("Data/Default.csv")

glimpse(dat)
tail(dat)
slice_sample(dat,n = 10)

# hot swap 'Yes' and 'No' in default in student columns to 1 and 0
dat <- dat |>
  mutate(
    default = if_else(default == "Yes", 1, 0),
    student = if_else(student == "Yes", 1, 0)
  )
```

## Q.5.A
```{r}
# fit a logistic a regression model to the training data
set.seed(1)
modelA <- glm(default ~ income + balance, family = "binomial", data = dat)
summary(modelA)
```

## Q.5.B
```{r}
## The validation set approach
set.seed(1)

# naively split the data into training and test sets with 80% training.
obs <- dat |> nrow()
train <- sample(obs, 0.8*obs)

# fit a logistic a regression model to the training data
modelB1 <- glm(default ~ income + balance, family = "binomial", data = dat, subset = train)
summary(modelB1)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelB1, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)

```

```{r}
set.seed(1)

# naively split the data into training and test sets with 75% training
obs <- dat |> nrow()
train <- sample(obs, 0.75*obs)

# fit a logistic a regression model to the training data
modelB2 <- glm(default ~ income + balance, family = "binomial", data = dat, subset = train)
summary(modelB2)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelB2, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)

```

```{r}
set.seed(1)

# naively split the data into training and test sets with 85% training
obs <- dat |> nrow()
train <- sample(obs, 0.85*obs)

# fit a logistic a regression model to the training data
modelB3 <- glm(default ~ income + balance, family = "binomial", data = dat, subset = train)
summary(modelB3)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelB3, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)

```
## Q.5.C

*There is essentially no change in the test error when decreasing the training size, but there is a noticeable change when increasing the training size.

## Q.5.D
```{r}
set.seed(1)

# naively split the data into training and test sets with 80% training
obs <- dat |> nrow()
train <- sample(obs, 0.8*obs)

# fit a logistic a regression model to the training data
modelD <- glm(default ~ income + balance + student, family = "binomial", data = dat, subset = train)
summary(modelD)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelD, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)
```

There is an increase the test error after including the student status. 


# 2. Question 10.A in Section 6.6 
```{r, include = FALSE}
library(faux)
```


```{r}
# generate a data set with p = 20 and n = 1000

set.seed(5)

X <- rnorm_multi(n = 1000, vars = 20, as.matrix = TRUE)
eps <- rnorm(1000, sd = 5)
beta <- rnorm(20)

# make some elements exactly zero in the beta vector
zero_indices <- c(1,3,7,9,13,15,17,19)
beta[zero_indices] <- 0

# calculate a random linear model
Y <- X %*% beta + eps

# combine y and x into a tibble
y_tib <- as_tibble(Y)
x_tib <- as_tibble(X)
dat <- bind_cols(y_tib, x_tib) |> rename(y = V1)

```

# Q.10.B
```{r}
# split the data into 100 training obs and 900 testing obs
set.seed(5)

obs <- dat |> nrow()
train_ind <- sample(obs, 100)

train <- dat[train_ind,]
test <- dat |> anti_join(train, by = "y")

```

# Q.10.C
```{r, include = FALSE}
library(leaps)
```

```{r}
# perform best subset selection on training
regfit_full_train <- regsubsets(y ~ ., train, nvmax = 20)
reg_sum_train <- summary(regfit_full_train)

MSE_train <- as_tibble(reg_sum_train$rss / nrow(train)) |> mutate(nv = 1:length(reg_sum_train$rss))
MSE_train |> ggplot(aes(y=value, x = nv)) +
  geom_point() +
  labs(x = "Number of Variables", y = "MSE", title = "Best Subset Selection on Training") +
  theme_bw()
  
```

```{r}
# perform best subset selection on test data
# 1. Prepare the test matrix (including an intercept column)
test_mat <- model.matrix(y ~ ., data = test)
test_mse <- rep(NA, 20)

# 2. Loop through each model size (1 to 20)
for (i in 1:20) {
  # Extract coefficients for the best model of size i
  coefi <- coef(regfit_full_train, id = i)
  
  # Predict Y for the test set using these coefficients
  # We subset test_mat to only include columns used in this specific model
  pred <- test_mat[, names(coefi)] %*% coefi
  
  # Calculate and store the Test MSE
  test_mse[i] <- mean((test$y - pred)^2)
}

# --- Plotting the Test MSE ---
MSE_test <- tibble(
  model_size = 1:20,
  test_mse = test_mse
)

MSE_test |> ggplot(aes(x = model_size, y = test_mse)) +
  geom_point() +
  labs(x = "Number of Variables", y = "MSE", title = "Best Subset Selection on Test") +
  theme_bw()

```
#Q.10.E
```{r}
min_mse <- MSE_test |> filter(test_mse == min(test_mse))
print(min_mse$model_size)
```

The test model takes has the lowest MSE with a model size of 6 variables.


# Q.10.F
```{r}
coef6 <- coef(regfit_full_train, id = 6)[-1]

coef0 <- beta[c(8,10,11,14,16,20)]

print(coef6)
print(coef0)
```
# Q.10.F
```{r}
diff <- coef0 - coef6
x_axis <- 1:6

diff_tib <- as_tibble(diff) |> add_column(x_axis)

diff_tib|> ggplot(aes(y =diff, x = x_axis)) +
  geom_point()
```


# 3. Question 11 in Section 6.6



