---
title: "HW4"
output: html_document
---
```{r}
library(tidyverse)
library(rsample)
```

# 1. Section 5.4 Question 5 Data Cleaning

```{r}
dat <- read.csv("Data/Default.csv")

glimpse(dat)

# hot swap 'Yes' and 'No' in default in student columns to 1 and 0
dat <- dat |>
  mutate(
    default = if_else(default == "Yes", 1, 0),
    student = if_else(student == "Yes", 1, 0)
  )
```

## Q.5.A

```{r}
# fit a logistic a regression model to the training data
set.seed(1)
modelA <- glm(default ~ income + balance, family = "binomial", data = dat)
summary(modelA)
```

## Q.5.B

```{r}
## The validation set approach
set.seed(1)

# naively split the data into training and test sets with 80% training.
obs <- dat |> nrow()
train <- sample(obs, 0.8*obs)

# fit a logistic a regression model to the training data
modelB1 <- glm(default ~ income + balance, family = "binomial", data = dat, subset = train)
summary(modelB1)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelB1, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)

```

```{r}
set.seed(1)

# naively split the data into training and test sets with 75% training
obs <- dat |> nrow()
train <- sample(obs, 0.75*obs)

# fit a logistic a regression model to the training data
modelB2 <- glm(default ~ income + balance, family = "binomial", data = dat, subset = train)
summary(modelB2)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelB2, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)

```

```{r}
set.seed(1)

# naively split the data into training and test sets with 85% training
obs <- dat |> nrow()
train <- sample(obs, 0.85*obs)

# fit a logistic a regression model to the training data
modelB3 <- glm(default ~ income + balance, family = "binomial", data = dat, subset = train)
summary(modelB3)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelB3, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)

```

## Q.5.C

**There is essentially no change in the test error when decreasing the training size, but there is a noticeable change when increasing the training size**

## Q.5.D

```{r}
set.seed(1)

# naively split the data into training and test sets with 80% training
obs <- dat |> nrow()
train <- sample(obs, 0.8*obs)

# fit a logistic a regression model to the training data
modelD <- glm(default ~ income + balance + student, family = "binomial", data = dat, subset = train)
summary(modelD)

## calculate the validation set error

# first get the predictions of the test set as probabilities
predictions <- predict(modelD, dat, type = "response")[-train]

# hot swap the probabilities to 1 and 0 using a threshold of 0.5
predictions <- if_else(predictions > 0.5, 1, 0)

misidentified <- sum(predictions != dat$default[-train])

setError <- misidentified / length(predictions) 

print(setError)
```

**There is an increase the test error after including the student status. **


# 2. Question 10.A in Section 6.6 

```{r}
library(faux)
```


```{r}
# generate a data set with p = 20 and n = 1000

set.seed(5)

X <- rnorm_multi(n = 1000, vars = 20, as.matrix = TRUE)
eps <- rnorm(1000, sd = 5)
beta <- rnorm(20)

# make some elements exactly zero in the beta vector
zero_indices <- c(1,3,7,9,13,15,17,19)
beta[zero_indices] <- 0

# calculate a random linear model
Y <- X %*% beta + eps

# combine y and x into a tibble
y_tib <- as_tibble(Y)
x_tib <- as_tibble(X)
dat <- bind_cols(y_tib, x_tib) |> rename(y = V1)

```

## Q.10.B

```{r}
# split the data into 100 training obs and 900 testing obs
set.seed(5)

obs <- dat |> nrow()
train_ind <- sample(obs, 100)

train <- dat[train_ind,]
test <- dat |> anti_join(train, by = "y")

```

## Q.10.C

```{r}
library(leaps)
```

```{r}
# perform best subset selection on training
regfit_full_train <- regsubsets(y ~ ., train, nvmax = 20)
reg_sum_train <- summary(regfit_full_train)

MSE_train <- as_tibble(reg_sum_train$rss / nrow(train)) |> mutate(nv = 1:length(reg_sum_train$rss))
MSE_train |> ggplot(aes(y=value, x = nv)) +
  geom_point() +
  labs(x = "Number of Variables", y = "MSE", title = "Best Subset Selection on Training") +
  theme_bw()
  
```

```{r}
# perform best subset selection on test data
# Prepare the test matrix (including an intercept column)
test_mat <- model.matrix(y ~ ., data = test)
test_mse <- rep(NA, 20)

# Loop through each model size (1 to 20)
for (i in 1:20) {
  
  # Extract coefficients for the best training model of size i
  coefi <- coef(regfit_full_train, id = i)
  
  # Predict Y for the test set using these coefficients
  pred <- test_mat[, names(coefi)] %*% coefi
  
  # Calculate and store the Test MSE
  test_mse[i] <- mean((test$y - pred)^2)
}

# combine model size and test MSE into a tibble for plotting
MSE_test <- tibble(
  model_size = 1:20,
  test_mse = test_mse
)

MSE_test |> ggplot(aes(x = model_size, y = test_mse)) +
  geom_point() +
  labs(x = "Number of Variables", y = "MSE", title = "Best Subset Selection on Test") +
  theme_bw()

```

## Q.10.E

```{r}
min_mse <- MSE_test |> filter(test_mse == min(test_mse))
print(min_mse$model_size)
```

**The test model takes has the lowest MSE with a model size of 6 variables.**


## Q.10.F

```{r}
# get the coefficients for the model with the lowest test MSE (model size 6). Ignore the intercept.
coef6 <- coef(regfit_full_train, id = 6)[-1]

coef0 <- beta[c(8,10,11,14,16,20)]

print(coef6)
print(coef0)
```

**There is a significant difference in some of the coefficients between the model with the lowest test MSE and the true beta. Two of the coefficients are quite similar.**

## Q.10.F

```{r}
p = 20

# intialize a vector to store the coefficient error for each model size
coef_error <- numeric(p)

for (r in 1:p) {
  # get the coefficients for the model with r predictors (including the intercept)
  coef_r <- coef(regfit_full_train, id = r)
  
  # Build full beta_hat vector of length p (zeros for excluded predictors)
  beta_hat <- numeric(p)
  
  pred_names <- names(coef_r)[names(coef_r) != "(Intercept)"]
  
  # Extract column indices from names like "X1" -> 1
  col_idx <- as.integer(sub("X", "", pred_names))
  beta_hat[col_idx] <- coef_r[pred_names]
  
  # calculate error
  coef_error[r] <- sqrt(sum((beta - beta_hat)^2))
}

# combine model size and coefficient error into a tibble for plotting
coef_error_df <- tibble(
  model_size = 1:p,
  coef_error = coef_error
)

coef_error_df |> ggplot(aes(x = model_size, y = coef_error)) +
  geom_point() +
  labs(x = "Model Size", y = "Coefficient Error", title = "Coefficient Error vs Model Size") +
  theme_bw()
```

**The coefficient error follows the MSE as the model size increases for the test set, with the lowest error at a model size of 6 variables.**

# 3. Question 11 in Section 6.6 Data Cleaning

```{r}
dat <- read.csv("Data/Boston.csv")
glimpse(dat)
```

```{r}
library(glmnet)
```


## Q.11.A

```{r}
# use lasso regression to predict crim
x <- model.matrix(crim ~ ., dat)[, -1] # exclude the intercept
y <- dat$crim

# combine into tibble
model_dat <- tibble(y = y, x)

# split the data into training and test sets. Use 50/50 split.
set.seed(1)
obs <- nrow(dat)
train_ind <- sample(obs, 0.5*obs)
train <- model_dat[train_ind, ]
test <- model_dat |> anti_join(train, by = "y")

# fit a lasso regression model to the training data
x_train <- as.matrix(train[, -1]) # exclude the response variable
y_train <- train$y
lasso_model <- glmnet(x_train, y_train, alpha = 1)
plot(lasso_model)

# use cross validation to find the best lambda
x_test <- as.matrix(test[, -1]) # exclude the response variable
y_test <- test$y
cv.out <- cv.glmnet(x_train, y_train, alpha = 1)
plot(cv.out)

# the best lambda
bestlam <- cv.out$lambda.min

# predict the test set using the best lambda
lasso.pred <- predict(lasso_model, s = bestlam,
newx = x_test)

# the mse of the test set prediction with the best lambda
best_lassoTest_mse <-mean((lasso.pred - y_test)^2)
best_lassoTest_mse

```

**Lasso had a test mse of 40.58 when using the best lambda value found using cross validation. The lasso model dropped two variables as well.**

```{r, include=FALSE}
library(pls)
```


```{r}
# use PCR to predict crim
set.seed(1)
pcr.fit <- pcr(y ~ ., data = train, scale = TRUE, validation = "CV")
summary(pcr.fit)

validationplot(pcr.fit, val.type = "MSEP")

# predict test set using all comps as it is the lowest MSE
pcr.pred <- predict(pcr.fit, test, ncomp = 13)
# the mse of the test set
best_pcrTest_mse <- mean((pcr.pred - y_test)^2)
best_pcrTest_mse

```

**PCR had a test MSE of 41.33 and shows that just 9 comps would explain almost all the variance in the data, but the test mse is lowest with 13 comps.**

## Q.11.B

```{r}
# Use the lambda value that gives the lowest test MSE to fit a lasso model on the full data set and report the non-zero coefficients.
out_lasso <- glmnet(x, y, alpha = 1)
lasso.coef <- predict(out_lasso, type = "coefficients",s = bestlam)[0:14,]
nonzero_coef <- lasso.coef[lasso.coef != 0]
print(nonzero_coef[-1]) # exclude the intercept
print(paste("The lasso test MSE:",best_lassoTest_mse))

out_pcr <- pcr(y ~ ., data = model_dat, scale = TRUE, validation = "CV", ncomp = 13)
pcr.coef <- coef(out_pcr, ncomp = 13)
print(pcr.coef)
print(paste("The pcr test MSE:",best_pcrTest_mse))

```

**The lasso test MSE is lower than the best PCR test MSE, but they are less than one unit apart. The coefficients differ significantly between the two models, but the lasso model dropped some variables. Due to lasso having a lower test MSE and a more interpretable model, I would say the the lasso model performs better than the PCR model.**

## Q.11.C

**Age and tax features are dropped in the lasso model, but all variables are included in the PCR model; therefore, the PCR technique proved no more useful than just using least squares.**
