---
title: "HW1"
author: " Jason R. May
output: html_document
---

```{R}
library(tidyverse)
```

# Question 1
```{r}

  # part a
  college <- read.csv("Data/College.csv")
  
  # part b
  rownames(college) <- college[, 1]
  View(college)
  college <- college[, -1]
  view(college)
  
  # part c
  
    ## part i.
    summary(college)
    
    ## part ii.
    pairs(college[,2:11])
    
    ## part iii.
    college |> 
      ggplot(aes(x = factor(Private), y = Outstate)) +
      geom_boxplot() +
      labs(title = "Out-of-State Tuition vs Private College",
           x = "Private College",
           y = "Out-of-State tuintion") +
      theme_bw()

    ## part iv.
    Elite <- rep("No", nrow(college))
    Elite[college$Top10perc > 50] <- "Yes"
    Elite <- as.factor(Elite)
    college <- data.frame(college, Elite)

    summary(college)

    college |>
      ggplot(aes(x = factor(Elite), y = Outstate)) +
      geom_boxplot() +
      labs(title = "Out-of-State Tuition vs Elite College",
           x = "Elite College",
           y = "Out-of-State tuintion") +
      theme_bw()

    ## part v.
    par(mfrow = c(2,2))
    bins = c(10, 15, 20, 25)
    for (bin in bins) {
      hist(college$Outstate, 
           breaks = bin, 
           main = paste("Histogram of Out-of-State Tuition with", bin, "bins"),
           xlab = "Out-of-State Tuition",
           col = "lightblue",
           border = "black")
    }

    ## part vi.

    # exploring the relation ship between grad rate and s.f.ratio
    college |>
      ggplot(aes(S.F.Ratio,Grad.Rate)) +
      geom_point() +
      labs(title = "Graduation Rate vs Student-Faculty Ratio",
           x = "Student-Faculty Ratio",
           y = "Graduation Rate") +
      theme_bw()

    # randomly sample half the data to meet the assumption of linear regression
    set.seed(100)
    collegeSample <- 
      college |> sample_n(nrow(college)/2)

    view(collegeSample)

    # run a linear model to see the relationship
    model <- 
      lm(Grad.Rate ~ S.F.Ratio, data = collegeSample)

    # check residuals
    plot(model)

    # The residuals appear to be linear, normal, and equally spread out; thus, the assumptions of linear regression are met.
    summary(model)
    
```
## Question 1 part C subpart VI summary
From a linear regression test, we found the assumption to be met and a significant negative relationship between Graduation Rate and Student-Faculty Ratio (t = -1.25, p-value < 0.001, df = 386). This suggests a higher student-faculty ratio is associated with a lower graduation rate. The overall model was statistically significant (F = 15.63, p-value < 0.001, df = 386) but explained approximately 9.4% of the variance in graduation rates (Adjusted R-squared: 0.094). Therefore, while there is a significant relationship, the effect size is small.

# Question 2
```{r}

```



