---
title: "HW5"
output: html_document
---
```{r, output = 'hide'}
library(tidyverse)
library(tidymodels)
library(splines)
library(mgcv)
library(tree)
library(randomForest)
library(ranger)
```



# 1. Question 9 in Section 7.9
```{r}
dat <- read_csv('Data/Boston.csv')
glimpse(dat)
```
### Q.1.A
```{r}
# fit a cubic regression model that predics nox given dis.
model <- lm(nox ~ poly(dis, 3), data = dat)

summary(model)

# plot the data and the fitted curve
ggplot(dat, aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ poly(x, 3), se = FALSE) +
  labs(title = "Cubic Regression of NOX on DIS",
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
```

### Q.1.B
```{r}
# plot the RSS over a range of degrees for the polynomial fit
rss_values <- sapply(1:15, function(degree) {
  model <- lm(nox ~ poly(dis, degree), data = dat)
  sum(model$residuals^2)
})

plot(1:15, rss_values, type = "b", xlab = "Degree of Polynomial", ylab = "RSS",
     main = "RSS vs Degree of Polynomial")
```
### Q.1.C
```{r, cahce = TRUE}
# use cv to find the best degree

# Cross-validation folds
set.seed(777)
folds <- vfold_cv(dat, v = 10)

# set the recipe 
rec <- recipe(nox ~ dis, data = dat) |>
  step_poly(dis, degree = tune())

# set the cooking method
mod <- linear_reg() |>
  set_engine("lm")

# set the plan
wf <- workflow() |>
  add_recipe(rec) |>
  add_model(mod)

# tune over a grid of degrees
degree_grid <- tibble(degree = 1:10)

tune_results <- tune_grid(
  wf,
  resamples = folds,
  grid      = degree_grid,
  metrics   = metric_set(rmse)
)

# pick the best degree
autoplot(tune_results)
show_best(tune_results, metric = "rmse")
best_degree <- select_best(tune_results, metric = "rmse")

#  fit on all of dat
final_wf  <- finalize_workflow(wf, best_degree)
final_fit <- fit(final_wf, data = dat)

# The optimal fit on a plot of the data
dat |> ggplot( aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ poly(x, best_degree$degree), se = FALSE) +
  labs(title = paste("Polynomial Regression of NOX on DIS (Degree =", best_degree$degree, ")"),
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
```

**We find the most optimal degree, based on minimal RMSE, is 3. However, it should be noted, that a degree 4 polynomial has essentially the same RMSE.**

### Q.1.D
```{r}
# fit a spline regression model that predicts nox given dis.
model_spline <- lm(nox ~ bs(dis, df = 4 ), data = dat )

summary(model_spline)

# plot the data and the fitted curve
dat |> ggplot(aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ bs(x, df = 4), se = FALSE) +
  labs(title = "Spline Regression of NOX on DIS",
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
                                                
```

**The default knots are used which are 25th, 50th, and 75th percentiles of dis.**

### Q.1.E
```{r}
# plot fitted spline RSS over range of degrees of freedom
rss_spline <- sapply(3:25, function(df) {
  model <- lm(nox ~ bs(dis, df = df), data = dat)
  sum(model$residuals^2)
})

plot(3:25, rss_spline, type = "b", xlab = "Degrees of Freedom", ylab = "RSS",
     main = "RSS vs Degrees of Freedom for Spline Regression")

```

**The most optimal df is likely 13, as the RSS only decreases slightly from then on out.**

### Q.1.F
```{r, cache = TRUE}
# Cross-validation folds
set.seed(777)
folds <- vfold_cv(dat, v = 10)

# set the recipe 
rec <- recipe(nox ~ dis, data = dat) |>
  step_bs(dis, deg_free = tune(), degree = tune())

# set the cooking method
mod <- linear_reg() |>
  set_engine("lm")

# set the plan
wf <- workflow() |>
  add_recipe(rec) |>
  add_model(mod)

# tune over a grid of df
df_grid <- grid_regular(
  deg_free(range = c(3, 12)),
  degree(range = c(1, 5)),
  levels = c(deg_free = 10, degree = 5)
)


tune_results <- tune_grid(
  wf,
  resamples = folds,
  grid      = df_grid,
  metrics   = metric_set(rmse)
)

# pick the best degree
autoplot(tune_results)
show_best(tune_results, metric = "rmse")
best_df <- select_best(tune_results, metric = "rmse")

#  fit on all of dat
final_wf  <- finalize_workflow(wf, best_df)
final_fit <- fit(final_wf, data = dat)

# The optimal fit on a plot of the data
dat |> ggplot(aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ bs(x, df = best_df$deg_free, knots = best_df$degree), se = FALSE) +
  labs(title = "Spline Regression of NOX on DIS",
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
```

**The optimal fit, defined as the lowest rmse, is df = 8, degree = 2. **

# 2. Question 10 in Section 7.9

```{r}
dat <- read_csv('Data/College.csv')
dat <-dat |>
  column_to_rownames('...1')
```

### Q.2.A
```{r}
set.seed(777)
split <- initial_split(dat, prop = 0.75)
train_data <- training(split)
test_data <- testing(split)

# perform forward step wise selection on the training data
full_model <- lm(Outstate ~ ., data = train_data)
null_model <- lm(Outstate ~ 1, data = train_data)

n <- nrow(mtcars)
forward_model <- MASS::stepAIC(null_model, scope = list(lower = null_model, upper = full_model), direction = "forward", k = log(n), trace = FALSE)

summary(forward_model)
```

### Q.2.B
```{r}
# fit GAM using predictors of the previous model
gam_model <- gam(Outstate ~ s(Expend) + Private + s(Room.Board) + s(Grad.Rate) + s(Terminal) + 
    s(perc.alumni) + s(Personal) , data = train_data)
summary(gam_model)

# plot the fitted functions
par(mfrow = c(2, 4))
plot(gam_model, se=TRUE, col = "blue")
```

**The GAM model resulted in a higher adjusted r-squared (79%) value over the forward step wise selection(75%). Though, this could just suggest over fitting. Another difference is that the terminal feature is now less significant.**

# Q.2.C
```{r}
# predict on the test data
pred_forward <- predict(forward_model, newdata = test_data)

pred_gam <- predict(gam_model, newdata = test_data)

rmse_forward <- sqrt(mean((test_data$Outstate - pred_forward)^2))
rmse_gam <- sqrt(mean((test_data$Outstate - pred_gam)^2))

cat("RMSE for Forward Stepwise Selection:", rmse_forward, "\n")
cat("RMSE for GAM:", rmse_gam, "\n")
```

**The gam model has a significantly lower RMSE on the test data, suggesting it is a better model. Therefore, some of the features are most likely non linear.**

# Q.2.D 

**Expend and Personal appear non-linear from the gam plots. **


# 3. Question 8 in Section 8.4
```{r}
dat <- read_csv('Data/Carseats.csv')

# convert character columns to numeric factors
dat <- dat |>
  mutate_if(is.character, as.factor) |>
  mutate_if(is.factor, as.numeric)
glimpse(dat)
```

### Q.3.A
```{r}
# split the data
set.seed(777)
split <- initial_split(dat, prop = 0.75)
train_data <- training(split)
test_data <- testing(split)
```


### Q.3.B
```{r}
# fit a regression tree
tree_model <- tree(Sales ~ ., data = train_data)

# print the summary of the tree model
summary(tree_model)

# plot the tree
plot(tree_model)
text(tree_model, pretty = 0)

# predict on the training
pred_train <- predict(tree_model, newdata = train_data)

# print training MSE
mse_train <- mean((train_data$Sales - pred_train)^2)
cat("\nTraining MSE for Regression Tree:", mse_train, "\n")

# predict on test
pred_tree <- predict(tree_model, newdata = test_data)

# print test MSE
mse_tree <- mean((test_data$Sales - pred_tree)^2)
cat("\nTest MSE for Regression Tree:", mse_tree, "\n")

```

**There is a significant increase in MSE from the training (2.33) to the test data (4.30), suggesting that the tree is over fitting.** 

### Q.3.C
```{r}
# prune the tree using cross validation
set.seed(777)

cv_tree <- cv.tree(tree_model)

# plot the cross validation results
plot(cv_tree$size, cv_tree$dev, type = "b", xlab = "Tree Size", ylab = "Deviance", main = "Cross-Validation for Tree Pruning")

# find the optimal tree size
optimal_size <- cv_tree$size[which.min(cv_tree$dev)]
cat("\nOptimal Tree Size:", optimal_size, "\n")

# prune the tree to the optimal size
pruned_tree <- prune.tree(tree_model, best = optimal_size)
# predict on the test data using the pruned tree
pred_pruned <- predict(pruned_tree, newdata = test_data)
# print test MSE for the pruned tree
mse_pruned <- mean((test_data$Sales - pred_pruned)^2)
cat("\nTest MSE for Pruned Tree:", mse_pruned, "\n")

```

**Pruning the tree would not help in reducing the test MSE as the minimum deviance is essentially at the largest tree size. Therefore, the tree is not over fitting due to its size. Even with pruning the test MSE is 4.44. **

### Q.3.D
```{r}
# fit a bagging model
set.seed(777)
bag_model <- randomForest(Sales ~ ., data = train_data, mtry = ncol(train_data) - 1, ntree = 500)

# predict on the test data and training
pred_bag_test <- predict(bag_model, newdata = test_data)
pred_bag_train <- predict(bag_model, newdata = train_data)

# print training MSE for the bagging model
mse_bag_train <- mean((train_data$Sales - pred_bag_train)^2)
cat("\nTraining MSE for Bagging Model:", mse_bag_train, "\n")

# print test MSE for the bagging model
mse_bag <- mean((test_data$Sales - pred_bag_test)^2)
cat("\nTest MSE for Bagging Model:", mse_bag, "\n")

# print the importance
randomForest::importance(bag_model)
```

**After implementing a basic bagging approach, the Test MSE has decreased but there is still significant over fitting considering the increase in MSE from training to test. The top 3 most important feature are: ShelveLoc, Price, and CompPrice. It should be noted, important features to a bad model are not necessarily important to a good one.**

### Q.3.E
```{r}

set.seed(777)

# use a random forest to predict sales, tune mtry, depth, and number of trees.

# set up the recipe
rec <- recipe(Sales ~ ., data = train_data)

# set up the model
rf_model <- rand_forest(
  mtry = tune(),
  trees = tune(),
  min_n = tune ()
) |>
  set_engine("ranger", max.depth = 5, num.threads = 1, importance = "impurity") |>
  set_mode("regression")

# set up the workflow
wf <- workflow() |>
  add_recipe(rec) |>
  add_model(rf_model)

cv_folds <- vfold_cv(train_data, v = 10)

n_features <- ncol(train_data) - 1  # Exclude the target variable
sqrt_features <- floor(sqrt(n_features))  # Common default for mtry

# set up the tuning grid
rf_grid <- grid_space_filling(
  mtry(range = c(sqrt_features, n_features)),
  trees(range = c(100, 500)),
  min_n(range = c(2, 20)),
  size = 30
)

```

```{r, cache = TRUE}
library(doParallel)
n_cores <- parallel::detectCores() - 1  # leave 1 core free for the OS
cl <- makeCluster(n_cores)
registerDoParallel(cl)

rf_res <- tune_grid(
  wf,
  resamples = cv_folds,
  grid      = rf_grid,
  metrics   = metric_set(yardstick::rmse, yardstick::rsq),
  control   = control_grid(verbose = FALSE, save_pred = TRUE, allow_par = TRUE)
)

stopCluster(cl)
registerDoSEQ()

autoplot(rf_res)                      # visualize hyperparameter landscape
show_best(rf_res, metric = "rmse")    # top 5 combos

best_params <- select_best(rf_res, metric = "rmse")

final_wf <- wf |>
  finalize_workflow(best_params)

final_fit <- final_wf |>
  fit(data = train_data)


```

```{r}
# add predictions to the training and test data
train_preds <- train_data |>
  bind_cols(predict(final_fit, train_data))

test_preds <- test_data |>
  bind_cols(predict(final_fit, test_data))

# print training and test MSE
train_mse <- mean((train_preds$Sales - train_preds$.pred)^2)
test_mse <- mean((test_preds$Sales - test_preds$.pred)^2)

cat("\nTraining MSE for Random Forest:", train_mse, "\n")
cat("\nTest MSE for Random Forest:", test_mse, "\n")
```

```{r}
final_fitted <- final_fit |> extract_fit_parsnip()
 
importance_tbl <- final_fitted |>
  vip::vi() |>
  arrange(desc(Importance))
 
cat("Feature Importances (Gini impurity):\n")
print(importance_tbl, n = 20)
```

**While there is still a decent increase in MSE between training and test data (1.52 -> 2.57), the increase and the MSE is overall significantly lower. Therefore, this tuned random forest has reduced over fitting while increasing accuracy. Additionally, the important features remain the same. The value of m is the length of the subset of features given to a single decision tree, it is crucial in reducing over fitting by making it harder for a single tree to memorize every data point. The increase in difficulty leads to each decision tree being weaker, but over all, the ensmble is more accurate with a lower MSE. **

### Q.3.F
```{r}
# use BART
library(BART)
set.seed(777)

xtrain <- train_data |> select(-Sales) |> as.matrix()
ytrain <- train_data$Sales

xtest <- test_data |> select(-Sales) |> as.matrix()
ytest <- test_data$Sales

bartfit <- gbart(xtrain, ytrain, x.test = xtest)

yhat.bart_train <- bartfit$yhat.train.mean
trainMSE <- mean((ytrain - yhat.bart_train)^2)

cat("\nTraining MSE for BART:", trainMSE, "\n")

yhat.bart <- bartfit$yhat.test.mean
testMSE <- mean((ytest - yhat.bart)^2)

cat("\nTest MSE for BART:", testMSE, "\n")

```
**The BART model has a training MSE of 0.586 and a test MSE of 1.210, which is less than the tuned random forest. Therefore, the BART model reduces over fitting while increasing accuracy more than the tuned random forest. Additionally, it is more computationally efficient. **

# Question 9 in Section 8.4
```{r}
dat <- read_csv('Data/OJ.csv')

# convert character columns to numeric factors
dat <- dat |>
  mutate_if(is.character, as.factor)

glimpse(dat)
```

### Q.4.A
```{r}
# split the data
set.seed(777)
split <- initial_split(dat, prop = 0.748)
train_data <- training(split)
test_data <- testing(split)
```

### Q.4.B
```{r}
# fit a classification tree
tree_model <- tree(Purchase ~ ., data = train_data)
summary(tree_model)
```

**The training error rate is 16% and the tree has 7 terminal nodes.**

### Q.4.C
```{r}
tree_model
```

**The tree predicsts MM if the LoyalCH value is less than 0.276.**

### Q.4.D
```{r}
plot(tree_model)
text(tree_model, pretty = 0)
```

**Here we see the series of decisions that leads to each prediction.**

### Q.4.E 
```{r}
# predict on the test data
pred_tree <- predict(tree_model, newdata = test_data, type = "class")

# print test error rate
test_error_rate <- mean(pred_tree != test_data$Purchase)
cat("\nTest Error Rate for Classification Tree:", test_error_rate, "\n")

# print the confusion matrix
table(pred_tree, test_data$Purchase)
```

### Q.4.F
```{r}
# prune the tree using cross validation
set.seed(777)
cv_tree <- cv.tree(tree_model)
```

### Q.4.G
```{r}
# plot the cross validation results
plot(cv_tree$size, cv_tree$dev, type = "b", xlab = "Tree Size", ylab = "Deviance", main = "Cross-Validation for Tree Pruning")
```

### Q.4.H
```{r}
# find the optimal tree size
optimal_size <- cv_tree$size[which.min(cv_tree$dev)]
cat("\nOptimal Tree Size:", optimal_size, "\n")
```

### Q.4.I
```{r}
# prune the tree to the optimal size
pruned_tree <- prune.tree(tree_model, best = optimal_size)
```

### Q.4.J
```{r}
# compare the training error rate between the pruned tree and the unpruned tree

# predict on the training data using the pruned tree
pred_pruned_train <- predict(pruned_tree, newdata = train_data, type = "class")

# print training error rate for the pruned tree
train_error_rate_pruned <- mean(pred_pruned_train != train_data$Purchase)

cat("\nTraining Error Rate for Unpruned Tree:", mean(pred_tree != train_data$Purchase), "\n")
cat("\nTraining Error Rate for Pruned Tree:", train_error_rate_pruned, "\n")

```

**The training error rate for the unpruned tree is higher. **

### Q.4.K
```{r}
# compare the test error rate of the pruned tree to the unpruned tree

# predict on the test data using the pruned tree
pred_pruned <- predict(pruned_tree, newdata = test_data, type = "class")

# print test error rate for the pruned tree
test_error_rate_pruned <- mean(pred_pruned != test_data$Purchase)

cat("\nTest Error Rate for Unpruned Tree:", test_error_rate, "\n")
cat("\nTest Error Rate for Pruned Tree:", test_error_rate_pruned, "\n ")


```

**The test error rate is higher for the Pruned tree. **
