---
title: "HW5"
output: html_document
---
```{r, output = 'hide'}
library(tidyverse)
library(tidymodels)
library(splines)
```



# Question 9 in Section 7.9
```{r}
dat <- read_csv('Data/Boston.csv')
glimpse(dat)
```
### Q.9.A
```{r}
# fit a cubic regression model that predics nox given dis.
model <- lm(nox ~ poly(dis, 3), data = dat)

summary(model)

# plot the data and the fitted curve
ggplot(dat, aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ poly(x, 3), se = FALSE) +
  labs(title = "Cubic Regression of NOX on DIS",
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
```

### Q.9.B
```{r}
# plot the RSS over a range of degrees for the polynomial fit
rss_values <- sapply(1:15, function(degree) {
  model <- lm(nox ~ poly(dis, degree), data = dat)
  sum(model$residuals^2)
})

plot(1:15, rss_values, type = "b", xlab = "Degree of Polynomial", ylab = "RSS",
     main = "RSS vs Degree of Polynomial")
```
### Q.9.C
```{r, cahce = TRUE}
# use cv to find the best degree

# Cross-validation folds
set.seed(777)
folds <- vfold_cv(dat, v = 10)

# set the recipe 
rec <- recipe(nox ~ dis, data = dat) |>
  step_poly(dis, degree = tune())

# set the cooking method
mod <- linear_reg() |>
  set_engine("lm")

# set the plan
wf <- workflow() |>
  add_recipe(rec) |>
  add_model(mod)

# tune over a grid of degrees
degree_grid <- tibble(degree = 1:10)

tune_results <- tune_grid(
  wf,
  resamples = folds,
  grid      = degree_grid,
  metrics   = metric_set(rmse)
)

# pick the best degree
autoplot(tune_results)
show_best(tune_results, metric = "rmse")
best_degree <- select_best(tune_results, metric = "rmse")

#  fit on all of dat
final_wf  <- finalize_workflow(wf, best_degree)
final_fit <- fit(final_wf, data = dat)

# The optimal fit on a plot of the data
dat |> ggplot( aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ poly(x, best_degree$degree), se = FALSE) +
  labs(title = paste("Polynomial Regression of NOX on DIS (Degree =", best_degree$degree, ")"),
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
```

**We find the most optimal degree, based on minimal RMSE, is 3. However, it should be noted, that a degree 4 polynomial has essentially the same RMSE.**

# Q.9.D
```{r}
# fit a spline regression model that predicts nox given dis.
model_spline <- lm(nox ~ bs(dis, df = 4 ), data = dat )

summary(model_spline)

# plot the data and the fitted curve
dat |> ggplot(aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ bs(x, df = 4), se = FALSE) +
  labs(title = "Spline Regression of NOX on DIS",
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
                                                
```

**The default knots are used which are 25th, 50th, and 75th percentiles of dis.**

```{r}
# plot fitted spline RSS over range of degrees of freedom
rss_spline <- sapply(3:25, function(df) {
  model <- lm(nox ~ bs(dis, df = df), data = dat)
  sum(model$residuals^2)
})

plot(3:25, rss_spline, type = "b", xlab = "Degrees of Freedom", ylab = "RSS",
     main = "RSS vs Degrees of Freedom for Spline Regression")

```

**The most optimal df is likely 13, as the RSS only decreases slightly from then on out.**

# Q.9.F
```{r, cache = TRUE}
# Cross-validation folds
set.seed(777)
folds <- vfold_cv(dat, v = 10)

# set the recipe 
rec <- recipe(nox ~ dis, data = dat) |>
  step_bs(dis, deg_free = tune(), degree = tune())

# set the cooking method
mod <- linear_reg() |>
  set_engine("lm")

# set the plan
wf <- workflow() |>
  add_recipe(rec) |>
  add_model(mod)

# tune over a grid of df
df_grid <- grid_regular(
  deg_free(range = c(3, 12)),
  degree(range = c(1, 5)),
  levels = c(deg_free = 10, degree = 5)
)


tune_results <- tune_grid(
  wf,
  resamples = folds,
  grid      = df_grid,
  metrics   = metric_set(rmse)
)

# pick the best degree
autoplot(tune_results)
show_best(tune_results, metric = "rmse")
best_df <- select_best(tune_results, metric = "rmse")

#  fit on all of dat
final_wf  <- finalize_workflow(wf, best_df)
final_fit <- fit(final_wf, data = dat)

# The optimal fit on a plot of the data
dat |> ggplot(aes(x = dis, y = nox)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y ~ bs(x, df = best_df$deg_free, knots = best_df$degree), se = FALSE) +
  labs(title = "Spline Regression of NOX on DIS",
       x = "Distance to Employment Centers (DIS)",
       y = "Nitric Oxides Concentration (NOX)") +
  theme_minimal()
```

**The optimal fit, defined as the lowest rmse, is df = 8, degree = 2. **

# Question 10 in Section 7.9


# Question 8 in Section 8.4
# Question 9 in Section 8.4







